Notes and Problems from My Research\vspace{-1ex}

### 2019-05-16 Hopf Algebra Notes/Hopf Algebras.tex
Haven't pushed in a while

%identity map
\DeclareMathOperator{\id}{id}

\DeclareMathOperator{\End}{End}

\newcommand*{\Hom}{\ensuremath\text{Hom}}

%Hopf algebra stuff
\newcommand*{\Vectk}{\operatorname{Vect}_k}
\newcommand*{\Algk}{\operatorname{Alg}_k}
\newcommand*{\Coalgk}{\operatorname{Coalg}_k}
\newcommand*{\calM}{\mathcal{M}}
\DeclareMathOperator{\Coalg}{Coalg}
\DeclareMathOperator{\Alg}{Alg}
\newcommand*{\scrT}{\ensuremath\mathscr{T}}
\newcommand*{\Ch}{\mathbf{Ch}}
\newcommand*{\g}{\mathfrak{g}}
\DeclareMathOperator{\Der}{Der}

@@

\begin{document}
%make the title page
\title{Hopf Algebras\vspace{-1ex}}
\author{A course by: Prof. James Zhang\\
Notes by: Nico Courts}
\date{Winter 2019}
\maketitle

@@

	This course is an introduction to Hopf algebras. In addition to basic material in 
	Hopf algebra, we will present some latest developments in quantum groups and tensor 
	and fusion categories. One of the newer topics is homological properties of 
	Noetherian Hopf algebras of low Gelfand-Kirillov dimension. A good reference for 
	the first two topics in the book \textit{Hopf Algebras and Their Action Rings} by 
	Susan Montgomery. Here is a list of possible topics:

\begin{itemize}
	\item Classical theorems concerning finite dimensional Hopf algebras.
	\item Infinite dimensional Hopf algebras and quantum groups.
	\item Duality and Calabi-Yau property.
	\item Actions of Hopf algebras and invariant theory.
	\item Representations of Hopf algebras, tensor and fusion categories.
\end{itemize}
\end{abstract}

\section{January 7, 2019}
If you don't know what a symmetric tensor category is, today is going to be a three 
star day. Max is 5.

>>>>

%Homological algebra stuff
\DeclareMathOperator{\id}{id}
\DeclareMathOperator{\End}{End}
\newcommand*{\Hom}{\ensuremath\text{Hom}}

@@

\begin{document}
%make the title page
\title{Homological Algebra\vspace{-1ex}}
\author{A course by: Prof. Julia Pevtsova\\
Notes by: Nico Courts}
\date{Spring 2019}
\maketitle

@@

	This course is an introductory course on homological algebra. We will be following the 
	book \textit{An Introduction to Homological Algebra} by Charles Weibel. We will be
	covering the following topics:

\begin{itemize}
	\item Chain complexes, homotopies, homology and long exact sequence in homology
	\item Resolutions, derived functors, Ext and Tor. Koszul complexes
	\item Group (co)homology
	\item Triangulated and derived categories
	\item Spectral sequences or open topic depending on the class interests
\end{itemize}
\end{abstract}

\section{April 1, 2019}
If you don't know what a symmetric tensor category is, today is going to be a three 
star day. Max is 5.

### 2019-05-16 Reading Notes/reading_notes.tex
Haven't pushed in a while

%categories
\newcommand*{\Mod}{\text{\textbf{Mod}}}
\newcommand*{\C}{\mathcal{C}}
\newcommand*{\T}{\mathcal{T}}
\let\O\relax
\newcommand*{\O}{\mathcal{O}}
\newcommand*{\Ass}{\text{\textbf{Ass}}}
\newcommand*{\Vect}{\text{\textbf{Vect}}}
\newcommand*{\Grp}{\text{\textbf{Grp}}}
\newcommand*{\Ab}{\text{\textbf{Ab}}}
\newcommand*{\Ring}{\text{\textbf{Ring}}}
\newcommand*{\Lie}{\text{\textbf{Lie}}}

@@

@@

\section{To be filed:}

>>>>

%categories
\newcommand*{\Mod}{\text{\textbf{Mod}}}
\newcommand*{\C}{\mathcal{C}}
\newcommand*{\T}{\mathcal{T}}
\let\O\relax
\newcommand*{\O}{\mathcal{O}}
\newcommand*{\Ass}{\text{\textbf{Ass}}}
\newcommand*{\Vect}{\text{\textbf{Vect}}}
\newcommand*{\Grp}{\text{\textbf{Grp}}}
\newcommand*{\Ab}{\text{\textbf{Ab}}}
\newcommand*{\Ring}{\text{\textbf{Ring}}}
\newcommand*{\Lie}{\text{\textbf{Lie}}}
\DeclareMathOperator{\stHom}{\underline{\Hom}}

@@

\part{Spring 2019}
\section{Meeting Notes}

\subsection{Week 3}
Due to travelling and some growing pains teaching my own 308 class research has gotten off to a slow
start. Luckily Julia's class requires us to do a presentation on homological algebra and last week 
we discussed some possible topics (mainly drawing on Dave Benson's course notes from Oberwolfach which I need 
to add as a reference here).

\begin{qst} Is the right way to think about triangles in a homotopy category where the suspension
measures failure to be exact?

Is a Cartesian square just a pushout/pullback ``modulo homotopy''? These seem to be related by 
the fact that a Cartesian square
\begin{center}
	\begin{tikzcd}
		A
	\end{tikzcd}
\end{center}
corresponds to a sequence
\[X\to Y'\sqcup Y''\to Z\to \Sigma X\]
which appears to be ``almost'' the fiber product except the $\gamma$ map.
\end{qst}

Maybe. Maybe not. Let's talk about it in the context of modular representation theory. The triangles 
here comes from short exact sequences $0\to M'\to M\to M''\to 0$. But a triangle in the 
triangluated category corresponds to $M'\to M\to M''\to \Sigma M$.

What this means is that 
\[\Ext_G^1(M'',M')\cong \stHom_G(M'',\Omega^{-1}M')\]

For next week, read Neeman's paper on triangulated categories and try reading through the computation
in \cite{julia-dave}.

\subsection{Week 4}
Most of my personal time has been spent on reading \cite{julia-dave}. I will summarize where I am at the moment.

I am reading theorem 4.1, which states that if $G=(\Ga^-\times \Ga^-)\rtimes \Ga_{(1)}$, then $H^{\ast,\ast}(G,k)$
is generated by $\zeta$ in degree $(1,1)$, $x\in(2,0)$, $\kappa\in(p,1)$ and $\lambda_i\in (i,i+1)$ for $1\le i\le p-2$.
All of these generators are found by computations earlier giving a presentation of $kG$ as a quotient of a polynomial ring:
\[\frac{k[u,v,t]}{(u^2,v^2,uv+vu,t^p, tv-vt, tu-ut-v)}\]

\begin{qst}
	We let $\mu_p$ be the (finite) group scheme of $p^{th}$ roots of unity and say that $\mu_p\times\mu_p$
	acts on $kG$ where the first element acts on $u$ and the second on $t$.
\end{qst}
Wait... is this literally just multiplication by a (principle? does that even necessarily exist in general?) $p^{th}$ root of unity?

\begin{qst}
	What does the ``fibre'' correspond to the the spectral sequence?
\end{qst}

\subsection{Week 5}
The main thing I have been considering this week has been the computation
\[d(\lambda^\varepsilon\eta^a\zeta^bx^c)\]
where all terms in the sum (using the Leibniz rule) die (since $\lambda,\zeta,x$ lie in the bottom row) except 
\[a\lambda^{\varepsilon+1}\eta^{a-1}\zeta^{b+1}x^c\]
where we've used that $d(\eta)=\lambda\zeta$. It is clear that for this to be nonzero, we need $a\le p-1$
and when $\varepsilon=1$. \textbf{It ends up there was an error}

\subsubsection{Elementary abelian group schemes}
$\Ga_{(r)}$ has group aglebra isomorphic to the the truncated polynomial algebra $k[x_i]/(x^p)$. Then an elementary abelian
group scheme is $\Ga_{(r)}\times(\bbZ/p)^k$. Now the homology (of the two factors?) are isomorphic, but
the Steenrod operations, which can further distinguish between them.

The Steenrod operations can bee seen as $\cal P$ and $\beta$, which is the connection homomorphism that arises 
from the long exact sequence in cohomology arising from $\bbZ/p\to\bbZ/p^2\to\bbZ/p$: $\beta:H^1(G,\mathbb{F}_p)\to H^2(G,\mathbb{F}_p)$. We can generalize to other
fields, but the $\beta$ map is no lnger linear, but $p$-semilinear!

There is a theorem by Serre on Bocksteins (?). Frattini quoitient -- maximal elementary abelian quotient -- look this up!

So the idea thien is we use this Frattini thing (think an analog to Nakayama) and we are trying to see when the Frattini quotient
is the group itself. 
\begin{lem}
	Let $\varphi:H\to G$ be a (surjective?) map of $p$-groups and $\varphi^*$ ist induced map on the holomology rings with trivial coefficients.
	then if $\varphi^*$ restricted to $H^1$ is an isomorphism and on $H^2$ is injectice, then $\varphi$ is an isomorphism.
\end{lem}
\begin{rmk}
	This requires a result that $H^1(G,k)\ne 0$ when $G$ is a $p$ group.
\end{rmk}

\subsection{Week 6}

\section{Group (Scheme) Cohomology}

To get some more background for reading \cite{julia-dave}, I began reading Evens' book on group cohomology \cite{evens}.
So I will begin here.

I need to get better about typing notes! It is so useful!

@@
